An Evolutionary Algorithm for the Multiobjective Risk-Equity Constrained Routing Problem

### 2011-11-30 DAW_hazmatHypE/trunk/paper/hazmatRouting_EvolOptim.tex
added text of W-HypE experiments

Instead of the environmental selection of HypE \citep{bz2011a}, other selection schemes can be easily used with the PISA framework and we show the resutls of two other algorithms exemplary. On the one hand, we employ the famous NSGA-II selection scheme \citep{dapm2002a} and, on the other hand, we use a generalized version of HypE towards a generalized weighted hypervolume indicator \citep{abbz2009b,bbtz2011a}. The boxplots of the resulting hypervolume indicator values after 100,000 function evaluations (population of size 100 after 1000 generations) are shown in Fig.~\ref{fig:algoComparison}. \TODO{What can be seen is that both? algorithms show a reduced performance regarding the standard hypervolume value. For NSGA-II, this is expected as its environmental selection schemes becomes less effective than in bi-objective problems when the number of objectives increases \citep{wbn2007a,bz2011a}. This is not the case for indicator-based algorithms such as HypE which were specifically proposed to circumvent the problems of algorithms optimizing non-compliant quality measures such as NSGA-II.

The reason for the reduced performance in the standard hypervolume indicator for W-Hype, however, is a different one. W-HypE was developed to steer the search towards regions of the objective space that have been specified by the user. To this end, it does not optimize the standard hypervolume indicator but a more generalized version of it, the weighted hypervolume indicator \citep{zbt2007a}. The idea behind W-HypE is that solutions in objective space regions with a higher weight will get a higher fitness due to the \emph{weighted} hypervolume they dominate. In order to be efficient also for a large number of objectives, W-HypE does not calculate the weighted hypervolume indicator values exactly (which is not possible for most of the weight functions) but instead samples them. Here, 10,000 samples are used in each iteration of the algorithm.

To show the principles of W-HypE in the application of our hazmat routing problem, we defined the weight of W-HypE's underlying weighted hypervolume indicator as constant in a box with corners $[7\cdot 10^3, 3\cdot 10^7, 5\cdot 10^5]$ and $[10^4, 4\cdot 10^7, 1.1\cdot 10^6]$ and zero otherwise. Instead of optimizing the standard hypervolume indicator as HypE, W-HypE now optimizes this weighted hypervolume indicator, resulting in better, i.e., dominating solutions in regions in or close to the box. Figure~\ref{fig:whype} shows the resulting non-dominated solutions of both algorithms with population size 25 after 10 independent runs of 4,000 generations each as well as the sampling box.}

This advantage of W-HypE over HypE within the pre-defined box can be observed both in early and late generations, cp.\ Fig.~\ref{fig:whype}.

\TODO{write}

>>>>

Instead of the environmental selection of HypE \citep{bz2011a}, other selection schemes can be easily used with the PISA framework and we show the resutls of two other algorithms exemplary. On the one hand, we employ the famous NSGA-II selection scheme \citep{dapm2002a} and, on the other hand, we use a generalized version of HypE towards a generalized weighted hypervolume indicator \citep{abbz2009b,bbtz2011a}. The boxplots of the resulting hypervolume indicator values after 100,000 function evaluations (population of size 100 after 1000 generations for NSGA-II and population size of 25 and 4,000 generations for W-HypE) are shown in Fig.~\ref{fig:algoComparison}. What can be seen is that both algorithms show a reduced performance regarding the standard hypervolume value with W-HypE being much worse than NSGA-II. For NSGA-II, this decrease  in hypervolume is expected as its environmental selection schemes does not optimize the hypervolume indicator as does HypE. Moreover, NSGA-II is known to become less effective than in bi-objective problems when the number of objectives increases \citep{wbn2007a,bz2011a}. This is not the case for indicator-based algorithms such as HypE which were specifically proposed to circumvent the problems of algorithms optimizing quality measures that do not comply with the Pareto dominance relation such as the crwoding distance of NSGA-II.

The reason for the reduced performance in the standard hypervolume indicator for W-HypE, on the other hand, can be explained in a different way. W-HypE and the underlying weighted hypervolume indicator was developed to steer the search towards regions of the objective space that have been specified by the user. To this end, instead of the standard hypervolume indicator, the weighted hypervolume indicator of \citep{zbt2007a} is optimized. The idea behind W-HypE is that solutions in objective space regions with a higher weight will get a higher fitness due to the \emph{weighted} hypervolume they dominate. Furthermore, in order to be efficient also for a large number of objectives, W-HypE does not calculate the weighted hypervolume indicator values exactly but instead samples them. Here, 10,000 samples are used in each iteration of the algorithm. For details on the algorithm, we refer to the original publication \citep{bz2011a}.

To show the principles of W-HypE in the application of our hazmat routing problem, we defined the weight of W-HypE's underlying weighted hypervolume indicator as constant in a box with corners $[7\cdot 10^3, 3\cdot 10^7, 5\cdot 10^5]$ and $[10^4, 4\cdot 10^7, 1.1\cdot 10^6]$ and zero otherwise. Instead of optimizing the standard hypervolume indicator as HypE does, W-HypE now optimizes this weighted hypervolume indicator as a secondary selection criterion after employing a non-dominated sorting. As a result, it is finding better, i.e., dominating solutions in regions in or close to the box. Figure~\ref{fig:whype} shows the resulting non-dominated solutions of both algorithms W-HypE and HypE with population size 25 after 10 independent runs of 4,000 generations together with the sampling box. While HypE originally finds 6 non-dominated solutions within the box, W-HypE finds 14 within the same budget of function evaluations. When both algorithm outcomes are combined, it turns out that 13 of W-HypE's solutions are still non-dominated while 5 out of the 6 solutions found by HypE are dominated by W-HypE solutions. It has to be noted that the advantage of W-HypE over HypE within the pre-defined box can be observed both in early and late generations (results not shown here).
